Searching for long-period transiting planets in the \kepler\ light curves using supervised classification

### 2015-03-30 document/ms.tex
some words and refs

\newcommand{\project}[1]{\textsl{#1}} % hogg say
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\KT}{\project{K2}}
\newcommand{\tess}{\project{TESS}}
\newcommand{\jwst}{\project{JWST}}
\newcommand{\terra}{\project{TERRA}}
\newcommand{\pdc}{\project{PDC}}
\newcommand{\license}{MIT License}
\newcommand{\projectname}{\project{ketu}}

@@

\title{%
    The search for long-period transiting planets in the \kepler\ light curves
}

\newcommand{\nyu}{2}
\newcommand{\mpia}{3}
\newcommand{\cds}{4}
\newcommand{\mpis}{5}
\author{%
    Daniel~Foreman-Mackey\altaffilmark{1,\nyu},
    David~W.~Hogg\altaffilmark{\nyu,\mpia,\cds},
    Bernhard~Sch\"olkopf\altaffilmark{\mpis},
    \etal
}
\altaffiltext{1}         {To whom correspondence should be addressed:
                          \url{danfm@nyu.edu}}
\altaffiltext{\nyu}      {Center for Cosmology and Particle Physics,
                          Department of Physics, New York University,
                          4 Washington Place, New York, NY, 10003, USA}
\altaffiltext{\mpia}     {Max-Planck-Institut f\"ur Astronomie,
                          K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{\cds}      {Center for Data Science, New York University,
                          726 Broadway, 7th Floor, New York, NY, 10003, USA}
\altaffiltext{\mpis}     {Max Planck Institute for Intelligent Systems
                          Spemannstrasse 38, 72076 T\"ubingen, Germany}

@@

Hi \citep{kois}.

@@

\clearpage
\bibliography{single}
\clearpage

>>>>

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\KT}{\project{K2}}
\newcommand{\tess}{\project{TESS}}
\newcommand{\jwst}{\project{JWST}}
\newcommand{\terra}{\project{TERRA}}
\newcommand{\pdc}{\project{PDC}}
\newcommand{\license}{MIT License}
\newcommand{\projectname}{\project{ketu}}

@@

\title{%
    Searching for long-period transiting planets in the \kepler\ light curves
    using supervised classification
}

\newcommand{\nyu}{2}
\newcommand{\cds}{3}
\newcommand{\mpia}{4}
\newcommand{\mpis}{5}
\author{%
    Daniel~Foreman-Mackey\altaffilmark{1,\nyu,\cds},
    David~W.~Hogg\altaffilmark{\nyu,\mpia,\cds},
    Bernhard~Sch\"olkopf\altaffilmark{\mpis},
    \etal
}
\altaffiltext{1}         {To whom correspondence should be addressed:
                          \url{danfm@nyu.edu}}
\altaffiltext{\nyu}      {Center for Cosmology and Particle Physics,
                          Department of Physics, New York University,
                          4 Washington Place, New York, NY, 10003, USA}
\altaffiltext{\cds}      {Center for Data Science, New York University,
                          726 Broadway, 7th Floor, New York, NY, 10003, USA}
\altaffiltext{\mpia}     {Max-Planck-Institut f\"ur Astronomie,
                          K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\altaffiltext{\mpis}     {Max Planck Institute for Intelligent Systems
                          Spemannstrasse 38, 72076 T\"ubingen, Germany}

@@

\begin{itemize}

\item Description of the physical significance of these planets

\item Discussion of previous attempts (papers by Gaudi, Yee, Payne, Bakos,
\etc).

\item Comparison of method to filtering methods and an argument for why
supervised classification should be more robust to false alarms given a large
enough training set

\end{itemize}

\citep{kois}

\section{Data preparation}

The \kepler\ Mission measured photometry for about 150,000 stars at half-hour
cadence for a baseline of over four years.
We aim to search these time series for single transits of long-period planets
and single eclipses of binary stars.
These data are made available on
MAST\footnote{\url{https://archive.stsci.edu/kepler/}} and, for each target,
we downloaded the full set of light curve files provided by Data Release 24
\citep{Thompson:2015}.
From these files, we extracted the PDC time series and split them into
``sections'' with no more than two contiguous missing or flagged data points.
The PDC light curves have been corrected for the instrumental effects caused
by the spacecraft using a data-driven model of the focal plane
\citep{Stumpe:2012, Smith:2012}.
Crucially, an attempt is also made by the PDC procedure to remove sharp
instrumental artifacts like ``sudden pixel sensitivity dropouts (SPSDs)''.
Any remaining missing data points are filled in using linear interpolation
resulting in sections of contiguous time series measurements with nearly
uniform time sampling.
This procedure yields about 150 \emph{light curve sections} for the typical
\kepler\ target.

The goal of this project is to discover the transits of long-period planets
that have not yet been discovered.
Therefore, when studying the light curve of an eclipsing binary star or a star
with known transiting planet candidates---on shorter periods---we also remove
all the in-transit data for the candidate using the parameters provided by the
\project{NASA Exoplanet
Archive}\footnote{\url{http://exoplanetarchive.ipac.caltech.edu/}; We
downloaded the \texttt{cumulative} table of \kepler\ Objects of Interest on
2015-03-25.}.

As we discuss further below, a key assumption of our analysis is that the
process generating any non-transit variability or noise in the light curve
is stationary.

Every three months, the \kepler\ spacecraft rolled 90~degrees to maintain its
orientation with respect to the Sun.
This means that each star landed on a different set of pixels each quarter
and, therefore, the systematic signals caused by any specific spacecraft
orientation won't be shared between neighboring quarters.
A year later, however, the pointing would return to nearly the same setup and
each star ends up back on many of the same pixels.
This manifests itself in the light curves by causing the observations
separated by a year to be qualitatively similar.

For reasons that will become clear in \sect{search}, we split the light curve
sections into 3 disjoint sets.
We use these sets to run a train/validate/test procedure so the goal is to
have representative samples in each set.
Therefore, we attempt to split the sections intelligently; evenly
distributing sections from the same quarter and season across the splits.
When this is not possible, we randomly assign the section to a split.

\section{Random forest classification}

A common task in the machine learning literature is called \emph{supervised
classification} where the goal is to separate objects into classes
represented by sets of labeled examples.
One example from the astronomy literature is the problem of star--galaxy
separation (\dfmtodo{cite}).
In this problem, there are a set of images or photometric measurements that
have been classified---by some other method---as either a star or galaxy
observation and the goal is to transfer these labels to a set of observations
that have not yet been classified.
For a more in-depth discussion of the application of these techniques in
astronomy, the interested reader is directed to \citet{Ivezic:2013}.

The Random Forest (RF) classification model \citep{Breiman:2001} has become
very popular in the machine learning community.
It has also been applied with great success in astronomy \citep[for
example]{Richards:2011, Richards:2012} \dfmtodo{others}.
The RF model works by building

\section{Search procedure}\sectlabel{search}

\paragraph{Generate the training \& validation sets}

\paragraph{Train}

\paragraph{Validate}

\paragraph{Test}

\section{Tuning parameters}

\paragraph{The parameters}

\paragraph{Objective function}

\section{Demonstration}

\paragraph{Successes}

\paragraph{Failures}

\section{Parameter estimation}

\section{Population inference}

\section{Results}

\section{Discussion}

@@

\clearpage
\bibliography{peerless}
\clearpage
